---
title: "Regression Model Assignment"
author: "Bodhi Goswami"
date: "10/27/2021"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

***Executive Summary***

This report was prepared to explore relations of Miles per Gallon (mpg) of
cars with respect to pertinent parameters a car might have. This included
weight of the car, gross horsepower, type of transmission (automatic or manual),
number of cylinders etc. The data set built in in R called *mtcars* was used to
obtain conclusions.

***Data set Overview***

The data set mtcars contains info from *1974 Motor Trend US magazine* and contains
performance chart of 32 vehicles from 1973 models. Below is a preview of mtcars.

```{r, echo=FALSE}
require(datasets); data("mtcars")
head(mtcars)
```

***Exploratory Analysis***

The coefficients used for Automatic/Manual transmission is *am* in the data above.
The values set are 0 for automatic and 1 for manual transmission. Upon a simple
boxplot shown in Figure 1, we can infer the mean milage for manual transmission
cars are higher than automatic transmission cars. The points shown in Purple are
the individual points plotted along with the boxplot for confirmation of the same.


***Model Selection for Inference on MPG v/s Transmission Type***

While Binary and Poisson models do not suit to be a good fit since this mpg is
a quantifiable outcome with an upper bound per this data. The two best suited
models, in my opinion, for selection were Liner Model and Generalized Linear Model.

A linear model was fit as shown below to infer impact on milage due to all the
vehicular parameters in the given dataset.

```{r}
fit<-lm(mpg~. -1,data=mtcars)
summary(fit) $coef
```

Under the Estimate column, this shows the estimated of impact each of these parameters
have on one unit increase in mpg. For instance, one unit increase in gross 
horsepower would result in 2% reduction in milage. Since this is an inverse relation, higher
number of milage can be achieved with lesser horsepower. On the contrary, higher number
of gears will result in higher milage per the summary output. Each additional
ton weight of car will reduce the Totale milage by about 4 miles/gallon. These are
very interesting facts obtained from the linear model.

As an attempt to fit liner model for mpg v/s our regressor of interest (transmission)
code below was used. Inferences follow.

```{r}
lm.fit<-lm(mpg~as.factor(am), data=mtcars)
mtcars$am<-as.factor(mtcars$am)
summary(lm.fit)$coef
```

This code sets am(0) which is automatic transmission and gives an Estimate mpg of 17.14.
Same can be observed in the boxplot on Figure 1. For the manual transmission the 
Estimated miage is 17.14+7.24, which is about 7.24 mpg more than automatic. Hence
from the linear model it can be inferred manual transmission cars serve higher milage
per the given data.

A generalized linear model fit was also performed to verify if the inference we obtained
is accurate.

```{r}
glm.fit<-glm(mpg~as.factor(am), data=mtcars)
summary(glm.fit)$coef
```

As expected, manual transmission cars are shown to have a higher milage of 7.24.
This proves parity between Linear and Generalized Linear Model for this dataset
to find mpg v/s transmission.

***Residual Plot and Diagonostics***

A resudual boxplot was done for mpg outcome upon the two regressors of transmissions.
It was obtained the the residuals are much higher for Manual Transmission compared
to Automatic Transmission. Please see plot shown on Apprendix Figure 2.

***Conclusion***

From the linear model discussed above it can be concluded that manual transmission
cars provide higher milage than automatic transmission cars. Approximate difference
in mpg for manual transmission cars was found to be around 7.2 miles per gallon.



\newpage

***Appendix***

```{r, echo = FALSE}
require(datasets); require(ggplot2)
mtcars$am<-as.factor(mtcars$am)
g = ggplot(mtcars, aes( x = as.factor(am), y = mpg))
g = g + geom_boxplot(aes(fill=as.factor(am)),alpha = 0.35)
g = g + geom_point(size =4, col = "purple", alpha = 0.75)
g = g + xlab("Transmission Type (0 for Automatic, 1 for Manual)") +
            ylab("Miles per Gallon (MPG)") + ggtitle("Mpg v/s Transmission")
g = g + theme(plot.title = element_text(hjust='0.5'))
g = g + labs(fill = "Transmission")
g
```

*Fig 1: Plot Showing Higher Average MPG values for Manual Transmission*

\newpage

```{r}
am.fit<-lm(mpg~am, data=mtcars)
plot(resid(am.fit)~mtcars$am,pch=19,col="lightblue", 
     xlab = "Transmission Type (0 for Automatic, 1 for Manual)", 
     ylab = "Residuals")
abline(h=0, col="red", lwd=3)
```

*Fig 2: BoxPlot of Residuals*
